Master's Project Journal

### 2019-04-17 thesis/NetworkImplemenation.tex
58.5

Figure \ref{fig:integer_lf} contains an \ac{RTL} diagram describing the implementation. Important to note is the shift applied to the result of the multiplication by \acs{kp} which, as the integration is performed with a larger width to avoid accumulator overflow, preserves the intended relationship between proportional and integral gains.
The Verilog implementation of the \ac{LF} using integer arithmetic can be found in \textsc{LoopFilter.v}, attached in Appendix \ref{adx:code} Listing \ref{lst:loop_filter}.

@@

\begin{figure}[h]
	\centering
	\subfloat[Vivado simulation.]{{\includegraphics[width=0.45\textwidth]{../sim_locking} }}\\
	\subfloat[Hardware measurement.]{{\includegraphics[width=0.45\textwidth]{../impl_locking} }}%
	\caption[\ac{ADPLL} 1 Vivado locking simulations]{\ac{ADPLL} 1 Vivado locking simulations.}
	\label{fig:sim_locking}
\end{figure}
In Figure \ref{fig:sim_locking} (a), the startup behaviour of the system can be observed. This simulation was performed with a reference at exactly $5~\si{\mega\hertz}$ and an initial deviation of $252~\si{\kilo\hertz}$, and the plot shows the inverse of the \ac{DCO} period against the time since the oscillator was enabled. After some initial oscillation the period settles around that of the reference. Once locked, cycle-to-cycle jitter was calculated to be $3.491~\si{\nano\second}$ using the standard deviation of the periods, with a worst case cycle-to-cycle change of $9.299~\si{\nano\second}$. Lock was visually determined to have occurred after $20.91~\si{\micro\second}$, after which the oscillator frequency began to change in a repetitive pattern. later the same test carried in hardware with identical initial conditions, the results of which are shown in Figure \ref{fig:sim_locking} (b). This plot is derived from data captured using an Agilent MSO7054A, using a setup that will be further explained in Section \ref{section:measurement_setup}. Despite lacking a non ideal reference, less violent initial oscillation, and a lesser until lock could be visually determined, the simulation provided a good insight into the system's behaviour once implemented on an \ac{FPGA}.

\subsubsection{\acs{ADPLL} 1 Design Summary}
The listed components were brought together in \textsc{NetworkADPLL.v}, attached in Appendix \ref{adx:code} Listing \ref{lst:network_adpll}. Important to note is that each \ac{ADPLL} only contains two \acl{PD}s despite during network operation the need for phase comparisons to be made with up to four neighbours. The two ``missing'' \ac{PD}s are in fact synthesised as part of the other \ac{ADPLL}s, with which the comparison would be made. This allows for a 50\% reduction in the number of \ac{PD}s required, while also ensuring identical error is received by both \ac{ADPLL}s. Each \ac{ADPLL} performs comparison with the \ac{DCO} ``above'' and ``left'' of them in the Cartesian grid, and a port in the module outputs the results for use by the other \ac{ADPLL}s involved. %TODO juvenile wording
Table \ref{table:adpll1} contains a brief overview of the bit widths, tuning ranges and other configuration information for this \ac{ADPLL}.
\begin{table}[!h]
	\begin{center}
		\begin{tabular}{|l|r|r|r|}
			\multicolumn{4}{c}{\ac{DCO} Tuning Range} \T\\
			\hline
			\multicolumn{1}{|c|}{-} & \multicolumn{1}{c|}{Minimum} & \multicolumn{1}{c|}{Step} & \multicolumn{1}{c|}{Maximum} \T\\
			\hline
			Frequency & $1.008~\si{\mega\hertz}$ & \multicolumn{1}{r|}{$62.988~\si{\kilo\hertz}$} & $8.944~\si{\mega\hertz}$ \T\\
			\hline
			Period & $992.0~\si{\nano\second}$ & \multicolumn{1}{r|}{$2.5~\si{\nano\second}$ (at $5~\si{\mega\hertz}$)} & $111.8~\si{\nano\second}$ \T\\
			\hline
		\end{tabular}
		\begin{tabular}{|l|r|l|r|}
			\multicolumn{4}{c}{Configuration as Implemented} \T\\
			\hline
			\ac{DCO} Counter Width & 12 bits & \ac{DCO} Control Width (max) & 12 bits \T\\
			\hline
			\ac{DCO} Bias Point & 79 & Runtime Gains & Enabled \T\\
			\hline
			Phase Detector Error Width & 7 & Error Sum Weight Width & 4 \T\\
			\hline
			\acs{kp} & $2^{-5}\times[0,15]\cap\mathbb{Z}$ & \acs{ki} & $2^{-8}\times[0,15]\cap\mathbb{Z}$ \T\\
			\hline
			Detection Resolution & $3.875~\si{\nano\second}$ & Detection Phase Resolution & $6.975\si{\degree}$ (at $5~\si{\mega\hertz}$)\\
			\hline
		\end{tabular}
	\end{center}
\caption[ADPLL Design 1 Summary]{ADPLL Design 1 Summary.}
\label{table:adpll1}
\end{table}

@@

Using this updated maximum number of inverters the frequency range can be recalculated. Firstly, the new minimum and maximum numbers of inverters are $309$ and $373$ respectively, with a mid point at $341$. The inverter delay and period step are then recomputed based on the centre frequency of $5~\si{\mega\hertz}$ as:
\begin{align}%TODO 2 x 2 x midpoint justification - in prev. chapter?
t_{step} &= \frac{T_{osc}}{0.5\times(n-\frac{k}{2})} = \frac{200\si{\nano\second}}{0.5\times 341} = 1.176~\si{\nano\second} \\
\tau_{inv} &= \frac{t_{step}}{2\times\text{inverters per step}} = \frac{t_{step}}{2\times2} = \frac{1.176~\si{\nano\second}}{4} = 293~\si{\pico\second}
\end{align}
The frequency range is then:
\begin{align} %TODO inconsisentl calculation approach with previous version
f_{osc} &= \frac{1}{T_{osc}} = \frac{1}{(n-2CC)\times 2\times\tau_{inv}} \\
f_{min} &= \frac{1}{T_{max}} = \frac{1}{(373-0)\times 2\times 293~\si{\pico\second}} = 4.571~\si{\mega\hertz} \\
f_{max} &= \frac{1}{T_{min}} = \frac{1}{(373-2\times32)\times 2\times 293~\si{\pico\second}} = 5.518~\si{\mega\hertz}
\end{align}

@@

\begin{table}[!h]%TODO redo numbers!
	\begin{center}
		\begin{tabular}{|l|r|r|r|}
			\multicolumn{4}{c}{\ac{DCO} Tuning Range} \T\\
			\hline
			\multicolumn{1}{|c|}{-} & \multicolumn{1}{c|}{Minimum} & \multicolumn{1}{c|}{Step} & \multicolumn{1}{c|}{Maximum} \T\\
			\hline
			Frequency & $4.571~\si{\mega\hertz}$ & $29.180~\si{\kilo\hertz}$ (at $5~\si{\mega\hertz}$) & $5.518~\si{\mega\hertz}$ \T\\
			\hline
			Period & $218.2~\si{\nano\second}$ & $1.176~\si{\nano\second}$ & $181.1~\si{\nano\second}$ \T\\
			\hline
		\end{tabular}
		\begin{tabular}{|l|r|l|r|}
			\multicolumn{4}{c}{Configuration as Implemented} \T\\
			\hline
			\ac{RO} num. of inverters (max) & 373 & \ac{RO} Control Width & 5 bits \T\\
			\hline
			\ac{RO} Bias Point & 16 & Runtime Gains & Enabled \T\\
			\hline
			Phase Detector Error Width & 5 & Error Sum Weight Width & 4 \T\\
			\hline
			\acs{kp} & $2^{-7}\times[0,15]\cap\mathbb{Z}$ & \acs{ki} & $2^{-9}\times[0,15]\cap\mathbb{Z}$ \T\\
			\hline
			Detection Resolution & $3.875~\si{\nano\second}$ & Detection Phase Resolution & $6.975\si{\degree}$ (at $5~\si{\mega\hertz}$)\\
			\hline
		\end{tabular}
	\end{center}
	\caption[\ac{ADPLL} Design 2 Summary]{\ac{ADPLL} Design 2 Summary.}
	\label{table:adpll2}
\end{table}

@@

\subsubsection{\acs{ADPLL} 3 Design Summary}
Table \ref{table:adpll3}.

\begin{table}[!h]%TODO numbers!
	\begin{center}
		\begin{tabular}{|l|r|r|r|}
			\multicolumn{4}{c}{\ac{DCO} Tuning Range} \T\\
			\hline
			\multicolumn{1}{|c|}{-} & \multicolumn{1}{c|}{Minimum} & \multicolumn{1}{c|}{Step} & \multicolumn{1}{c|}{Maximum} \T\\
			\hline
			Frequency & $4.571~\si{\mega\hertz}$ & $29.180~\si{\kilo\hertz}$ (at $5~\si{\mega\hertz}$) & $5.518~\si{\mega\hertz}$ \T\\
			\hline
			Period & $218.2~\si{\nano\second}$ & $1.176~\si{\nano\second}$ & $181.1~\si{\nano\second}$ \T\\
			\hline
		\end{tabular}
		\begin{tabular}{|l|r|l|r|}
			\multicolumn{4}{c}{Configuration as Implemented} \T\\
			\hline
			\ac{RO} num. of inverters (max) & 373 & \ac{RO} Control Width & 5 bits \T\\
			\hline
			\ac{RO} Bias Point & 16 & Runtime Gains & Enabled \T\\
			\hline
			Phase Detector Error Width & 5 & Error Sum Weight Width & 4 \T\\
			\hline
			\acs{kp} & $2^{-7}\times[0,15]\cap\mathbb{Z}$ & \acs{ki} & $2^{-9}\times[0,15]\cap\mathbb{Z}$ \T\\
			\hline
			Detection Resolution & $3.875~\si{\nano\second}$ & Detection Phase Resolution & $6.975\si{\degree}$ (at $5~\si{\mega\hertz}$)\\
			\hline
		\end{tabular}
	\end{center}
	\caption[\ac{ADPLL} Design 3 Summary]{\ac{ADPLL} Design 3 Summary.}
	\label{table:adpll3}
\end{table}

@@

>>>>

Figure \ref{fig:integer_lf} contains an \ac{RTL} diagram describing the implementation. Important to note is the shift applied to the result of the multiplication by \acs{kp} which, as the integration is performed with a larger width to avoid accumulator overflow, preserves the intended relationship between proportional and integral gains. The value of $k_p$ and $k_i$ can be computed using the following formulae:
\begin{gather}
\text{Proportional gain} = k_p\times 2^{-path\_width\_change\_p} \\
path\_width\_change\_p = lsh\_width - rsh\_width \\
lsh\_width = ki\_width - kp\_width  \\
rsh\_width = error\_width + kp\_width + lsh\_width - control\_code\_width \\
\text{Integral gain} = k_i\times 2^{-path\_width\_change\_i} \\
path\_width\_change\_i = - rsh\_width
\end{gather}
In the example system shown in Figure \ref{fig:integer_lf} the proportional gain can be computed as $k_p\times 2^{3-8} = \frac{k_p}{32}$ and the integral gain is $k_i\times 2^{-8} = \frac{k_i}{128}$.
The Verilog implementation of the \ac{LF} using integer arithmetic can be found in \textsc{LoopFilter.v}, attached in Appendix \ref{adx:code} Listing \ref{lst:loop_filter}.

@@

\begin{figure}[h]
	\centering
	\subfloat[Vivado simulation.]{{\includegraphics[width=0.45\textwidth]{../sim_locking} }}%
	\subfloat[Hardware measurement.]{{\includegraphics[width=0.45\textwidth]{../impl_locking} }}%
	\caption[\ac{ADPLL} 1 Vivado locking simulations]{\ac{ADPLL} 1 Vivado locking simulations.}
	\label{fig:sim_locking}
\end{figure}
In Figure \ref{fig:sim_locking} (a), the startup behaviour of the system can be observed. This simulation was performed with a reference at exactly $5~\si{\mega\hertz}$ and an initial deviation of $252~\si{\kilo\hertz}$, and the plot shows the inverse of the \ac{DCO} period against the time since the oscillator was enabled. After some initial oscillation the period settles around that of the reference. Once locked, cycle-to-cycle jitter was calculated to be $3.491~\si{\nano\second}$ using the standard deviation of the periods, with a worst case cycle-to-cycle change of $9.299~\si{\nano\second}$. Lock was visually determined to have occurred after $20.91~\si{\micro\second}$, after which the oscillator frequency began to change in a repetitive pattern. later the same test carried in hardware with identical initial conditions, the results of which are shown in Figure \ref{fig:sim_locking} (b). This plot is derived from data captured using an Agilent MSO7054A, using a setup that will be further explained in Section \ref{section:measurement_setup}. Despite lacking a non ideal reference, less violent initial oscillation, and a lesser until lock could be visually determined, the simulation provided a good insight into the system's behaviour once implemented on an \ac{FPGA}.

\subsubsection{\acs{ADPLL} 1 Design Summary}
The listed components were brought together in \textsc{NetworkADPLL.v}, attached in Appendix \ref{adx:code} Listing \ref{lst:network_adpll}. Important to note is that each \ac{ADPLL} only contains two \acl{PD}s despite during network operation the need for phase comparisons to be made with up to four neighbours. The two ``missing'' \ac{PD}s are in fact synthesised as part of the other \ac{ADPLL}s, with which the comparison would be made. This allows for a 50\% reduction in the number of \ac{PD}s required, while also ensuring identical error is received by both \ac{ADPLL}s. Each \ac{ADPLL} performs comparison with the \ac{DCO} ``above'' and ``left'' of them in the Cartesian grid, and a port in the module outputs the results for use by the other \ac{ADPLL}s involved. %TODO juvenile wording
Table \ref{table:adpll1} contains a brief overview of the bit widths, tuning ranges and other configuration information for this \ac{ADPLL}.
\begin{table}[!h]
	\begin{center}
		\begin{tabular}{|l|r|r|r|}
			\multicolumn{4}{c}{\ac{DCO} Tuning Range} \T\\
			\hline
			\multicolumn{1}{|c|}{-} & \multicolumn{1}{c|}{Minimum} & \multicolumn{1}{c|}{Step} & \multicolumn{1}{c|}{Maximum} \T\\
			\hline
			Frequency & $1.008~\si{\mega\hertz}$ & \multicolumn{1}{r|}{$62.988~\si{\kilo\hertz}$} & $8.944~\si{\mega\hertz}$ \T\\
			\hline
			Period & $992.0~\si{\nano\second}$ & \multicolumn{1}{r|}{$2.5~\si{\nano\second}$ (at $5~\si{\mega\hertz}$)} & $111.8~\si{\nano\second}$ \T\\
			\hline
		\end{tabular}
		\begin{tabular}{|l|r|l|r|}
			\multicolumn{4}{c}{Configuration as Implemented} \T\\
			\hline
			\ac{DCO} Counter Width & 12 bits & \ac{DCO} Control Width (max) & 12 bits \T\\
			\hline
			\ac{DCO} Bias Point & 79 & Runtime Gains & Enabled \T\\
			\hline
			Phase Detector Error Width & 7 & Error Sum Weight Width & 4 \T\\
			\hline
			\acs{kp} width & $2^{-5}\times[0,15]\cap\mathbb{Z}$ & \acs{ki} & $2^{-8}\times[0,15]\cap\mathbb{Z}$ \T\\
			\acs{kp} & $2^{-5}\times[0,15]\cap\mathbb{Z}$ & \acs{ki} & $2^{-8}\times[0,15]\cap\mathbb{Z}$ \T\\
			\hline
			Detection Resolution & $3.875~\si{\nano\second}$ & Detection Phase Resolution & $6.975\si{\degree}$ (at $5~\si{\mega\hertz}$)\\
			\hline
		\end{tabular}
	\end{center}
\caption[ADPLL Design 1 Summary]{ADPLL Design 1 Summary.}
\label{table:adpll1}
\end{table}

@@

Using this updated maximum number of inverters the frequency range can be recalculated. Firstly, the new minimum and maximum numbers of inverters are $309$ and $373$ respectively, with a mid point at $341$. The inverter delay and period step are then recomputed based on the centre frequency of $5~\si{\mega\hertz}$ as:
\begin{align}%TODO 2 x 2 x midpoint justification - in prev. chapter?
t_{step} &= \frac{T_{osc}}{0.5\times(n-\frac{k}{2})} = \frac{200\si{\nano\second}}{0.5\times 341} = 1.176~\si{\nano\second} \\
\tau_{inv} &= \frac{t_{step}}{2\times\text{inverters per step}} = \frac{t_{step}}{2\times2} = \frac{1.176~\si{\nano\second}}{4} = 293~\si{\pico\second}
\end{align}
The frequency range is then:
\begin{align} 
f_{osc} &= \frac{1}{T_{osc}} = \frac{1}{(n-2CC)\times 2\times\tau_{inv}} \\
f_{min} &= \frac{1}{T_{max}} = \frac{1}{(373-0)\times 2\times 293~\si{\pico\second}} = 4.571~\si{\mega\hertz} \\
f_{max} &= \frac{1}{T_{min}} = \frac{1}{(373-2\times32)\times 2\times 293~\si{\pico\second}} = 5.518~\si{\mega\hertz}
\end{align}

@@

\begin{table}[!h]
	\begin{center}
		\begin{tabular}{|l|r|r|r|}
			\multicolumn{4}{c}{\ac{DCO} Tuning Range} \T\\
			\hline
			\multicolumn{1}{|c|}{-} & \multicolumn{1}{c|}{Minimum} & \multicolumn{1}{c|}{Step} & \multicolumn{1}{c|}{Maximum} \T\\
			\hline
			Frequency & $4.571~\si{\mega\hertz}$ & $29.180~\si{\kilo\hertz}$ (at $5~\si{\mega\hertz}$) & $5.518~\si{\mega\hertz}$ \T\\
			\hline
			Period & $218.2~\si{\nano\second}$ & $1.176~\si{\nano\second}$ & $181.1~\si{\nano\second}$ \T\\
			\hline
		\end{tabular}
		\begin{tabular}{|l|r|l|r|}
			\multicolumn{4}{c}{Configuration as Implemented} \T\\
			\hline
			\ac{RO} num. of inverters (max) & 373 & \ac{RO} Control Width & 5 bits \T\\
			\hline
			\ac{RO} Bias Point & 16 & Runtime Gains & Enabled \T\\
			\hline
			Phase Detector Error Width & 5 & Error Sum Weight Width & 4 \T\\
			\hline
			\acs{kp} width & $4$ & \acs{ki} & $6$ \T\\
			\hline
			\acs{kp} & $2^{-4}\times[0,15]\cap\mathbb{Z}$ & \acs{ki} & $2^{-6}\times[0,15]\cap\mathbb{Z}$ \T\\
			\hline
			Detection Resolution & $3.875~\si{\nano\second}$ & Detection Phase Resolution & $6.975\si{\degree}$ (at $5~\si{\mega\hertz}$)\\
			\hline
		\end{tabular}
	\end{center}
	\caption[\ac{ADPLL} Design 2 Summary]{\ac{ADPLL} Design 2 Summary.}
	\label{table:adpll2}
\end{table}

@@

\subsubsection{\acs{ADPLL} 3 Design Summary}
Compared to the oscillators suggested earlier, \ac{ADPLL} 3 achieves a significantly finer detection resolution, at the expense of significant variation in the detection step. This can be viewed as advantageous however, as the \ac{ADPLL} exhibits an exaggerated version of the implementation and manufacturing defect based variability present in an \ac{ASIC}. It is this type of \ac{ADPLL} that both the \acs{UCD} and Sorbonne research teams are using to examine potential designs and test theoretical results. As before, Table \ref{table:adpll3} presents the important numeric characteristics of this \ac{ADPLL}.

\begin{table}[!h]
	\begin{center}
		\begin{tabular}{|l|r|r|r|}
			\multicolumn{4}{c}{\ac{DCO} Tuning Range} \T\\
			\hline
			\multicolumn{1}{|c|}{-} & \multicolumn{1}{c|}{Minimum} & \multicolumn{1}{c|}{Step} & \multicolumn{1}{c|}{Maximum} \T\\
			\hline
			Frequency & $4.571~\si{\mega\hertz}$ & $29.180~\si{\kilo\hertz}$ (at $5~\si{\mega\hertz}$) & $5.518~\si{\mega\hertz}$ \T\\
			\hline
			Period & $218.2~\si{\nano\second}$ & $1.176~\si{\nano\second}$ & $181.1~\si{\nano\second}$ \T\\
			\hline
		\end{tabular}
		\begin{tabular}{|l|r|l|r|}
			\multicolumn{4}{c}{Configuration as Implemented} \T\\
			\hline
			\ac{RO} num. of inverters (max) & 373 & \ac{RO} Control Width & 5 bits \T\\
			\hline
			\ac{RO} Bias Point & 16 & Runtime Gains & Enabled \T\\
			\hline
			Phase Detector Error Width & 5 & Error Sum Weight Width & 4 \T\\
			\hline
			\acs{kp} width & $4$ & \acs{ki} & $6$ \T\\
			\hline
			\acs{kp} & $2^{-4}\times[0,15]\cap\mathbb{Z}$ & \acs{ki} & $2^{-6}\times[0,15]\cap\mathbb{Z}$ \T\\
			\hline
			Detection Resolution & $\approx0.63~\si{\nano\second}$ & Detection Phase Resolution & $1.134\si{\degree}$ (at $5~\si{\mega\hertz}$)\\
			\hline
		\end{tabular}
	\end{center}
	\caption[\ac{ADPLL} Design 3 Summary]{\ac{ADPLL} Design 3 Summary.}
	\label{table:adpll3}
\end{table}

@@

% //9 for older tests @1111 = 4@0000, but >> 5 cos of extra size
% <4 is a no can do //7 for older tests @1111 = 4@0001
